---
title: "DATA607Assignment Tidyverse"
author: "Dhanya Nair"
date: "2024-03-15"
output: 
  html_document: default
  pdf_document: default
---
## Data from fivethirtyeight website
Source: recent grads list downloaded from the site: [via]https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/

readme file to understand the variables: [via]https://github.com/fivethirtyeight/data/blob/master/college-majors/readme.md

## Packages being used are purr,dplyr, stringr, forcats within the tidyverse package

## Load packages

```{r,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(purrr)
library(stringr)
library(forcats) 
```


## Reading data from the source file location in fivethirtyeight college-majors grads dataset

```{r, warning=FALSE}
college_grads <- read.csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/recent-grads.csv')

str(college_grads)
```

## Using purr package functions for data analysis and transformation
```{r}

# using map() function in purr package
# map()returns a a vector or  multiple vectors based on the input provided
# it performs the function for all the values in the vector.It performs sort function in this case.

college_majors_final <- college_grads  %>%
  count(Major_category) %>%
  rename(no_of_majors =n) %>%
  map(sort, decreasing = TRUE)

# 2 vectors major category and no. of majors are returned
head(college_majors_final)


# map_chr() always returns a character vector and we can perform character functions
college_majors_final$Major_category <- college_majors_final$Major_category  |> 
  map_chr(str_replace_all, "&.+","")


# example using discard()
college_majors_final$no_of_majors <- college_majors_final$no_of_majors |>
  discard(~.<10)

# the vectors ost transformation
head(college_majors_final)

```

## Q: What is the college major category distribution by men and women?
## A: Use ggplot to compare various major category populations

```{r}
#summarize to a table aggregating major category, count of men and count of women variable
major_category <- college_grads  %>%
  group_by(Major_category) %>%
  summarize(count_Men = sum(Men,na.rm = TRUE),count_women = sum(Women,na.rm = TRUE) ) 

#convert to a data frame
major_category_wide <- as.data.frame(major_category )


#pivot the dataframe to longer
major_category_long <- pivot_longer(major_category_wide , cols = c(count_Men, count_women), names_to = "gender", values_to = "count")

# remove "count" text  from the gender variable values
major_category_long$gender <- str_replace_all(major_category_long$gender,".*_","")


# Scale count values to thousands
major_category_long$count <- major_category_long$count / 1000


#visualize the data
# show male and female count side by side using position = dodge
# angle the x axis test by 45
#change the color of the fill variable using scale_fill_manual function
major_category_long  %>% 
  ggplot(aes(x = Major_category , y = count , fill = gender)) + 
  geom_col(position="dodge") +
  labs(title = "Counts of Men and Women by College Major Category",
       x = "College Major Category",
       y = "Count(in thousands)",
       fill = "Gender") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("blue", "pink"))


```

## How do I flip coordinates?

```{r}
major_category_long  %>% 
  ggplot(aes(x = Major_category , y = count , fill = gender)) + 
  geom_col(position="dodge") +
  labs(title = "Counts of Men and Women by College Major Category",
       x = "College Major Category",
       y = "Count(in thousands)",
       fill = "Gender") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("blue", "pink")) +coord_flip()

```


## How do I change sort order to view the most count values first?
i used the fct_reorder function in forcats package


```{r}

major_category_long  %>% 
  ggplot(aes(x = fct_reorder(Major_category,count)  , y = count , fill = gender)) + 
  geom_col(position="dodge") +
  labs(title = "Counts of Men and Women by College Major Category",
       x = "College Major Category",
       y = "Count(in thousands)",
       fill = "Gender") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("blue", "pink")) +coord_flip()

```
