---
title: 'tidyverse: using forcats to improve your ggplots'
author: "catlin"
date: "2/10/2021"
output:
  pdf_document: default
  html_document: default
---

Note that although forcats is part of the "tidyverse", it is not automatically loaded when you run `library(tidyverse)`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(forcats)
```

# Handy forcats functions for ggplot2

## Comparing followers of world religions

Source: https://en.wikipedia.org/wiki/List_of_religious_populations

I was looking for a simple dataset with count data for many items to demonstrate some basic forcats functions that are useful when creating plots.

```{r, warning=FALSE}
religions = read_csv("https://raw.githubusercontent.com/acatlin/data/master/religions.csv", 
                  col_names = FALSE) %>% 
    rename(religion = X1, followers = X2) %>% 
    mutate(millions_of_followers = followers/1000000.0) %>% 
  select(religion, millions_of_followers)

religions
```

## 1A: basic ggplot

Q: What are the most followed religions?
A: Use ggplot to compare religious populations

You can also embed plots, for example:

```{r}
religions %>% 
  ggplot(aes(x = religion, y = millions_of_followers)) + 
    geom_col(fill = "lightblue") + 
    labs(x = "religion", y = "millions of followers", 
           caption = "https://en.wikipedia.org/wiki/List_of_religious_populations") 
```
 
## 1B: How do I flip coordinates?

```{r}
religions %>% 
  ggplot(aes(x = religion, y = millions_of_followers)) + 
    geom_col(fill = "lightblue") + 
    labs(x = "religion", y = "millions of followers", 
           caption = "https://en.wikipedia.org/wiki/List_of_religious_populations") + coord_flip()
```

## 2A: How do I change sort order?
Revised by: Andy Catlin

Q: How do we change the chart to show the most followed religions first?
A: Use forcats::fct_reorder()

```{r}
library(forcats) 

ggplot(religions, aes(x = fct_reorder(religion, millions_of_followers), 
                      y = millions_of_followers)) + 
  geom_col(fill = "lightblue") + 
  labs(x = "religion", y = "millions of followers", 
      caption = "https://en.wikipedia.org/wiki/List_of_religious_populations") + 
  coord_flip()
```

## 2B: How do I combine less frequently used categories?

Q: How do we combine the less-followed religions into a single group?
A: Use forcats::fct_other()

```{r}
top5 = unlist(select(head(arrange(religions, desc(millions_of_followers)), 5), religion))

religions %>% 
  mutate(religion = fct_other(religion, keep = top5, other_level = "Other religions")) %>%     
  ggplot(aes(x = fct_reorder(religion, millions_of_followers), y = millions_of_followers)) +
    geom_col(fill = "lightblue") + 
    labs(x = "religion", y = "millions of followers", 
        caption = "https://en.wikipedia.org/wiki/List_of_religious_populations") + 
        coord_flip()
```

## 2C: Adding a title
Reference: https://www.geeksforgeeks.org/ggplot2-title-and-subtitle-with-different-size-and-color-in-r/ 

```{r}
religions %>% 
  mutate(religion = fct_other(religion, keep = top5, other_level = "Other religions")) %>%     
  ggplot(aes(x = fct_reorder(religion, millions_of_followers), y = millions_of_followers)) +
    geom_col(fill = "lightblue") + 
    labs(x = "religion", y = "millions of followers", 
         title = "Most Popular Religions",
         subtitle = "[2021]",
         caption = "https://en.wikipedia.org/wiki/List_of_religious_populations") +
    theme(plot.title = element_text(size = 18, color = "blue"),
        plot.subtitle = element_text(size = 14, color = "gold")) +
          coord_flip()
```


## Koohyar Code:
This plots a simple geometric fractal called the Koch snowflake. It's created by recursively replacing line segments with smaller line segments at certain angles. It is copied over code, not written by me. 

```{r KoohPy_code, echo=TRUE}


# Function to recursively generate Koch curve
koch_curve <- function(points, depth) {
  if (depth == 0) {
    return(points)
  } else {
    new_points <- matrix(NA, ncol = 2, nrow = nrow(points) * 4 - 3)
    new_points[1, ] <- points[1, ]
    new_points[nrow(new_points), ] <- points[nrow(points), ]
    for (i in 1:(nrow(points) - 1)) {
      p1 <- points[i, ]
      p2 <- points[i + 1, ]
      p3 <- (2 * p1 + p2) / 3
      p5 <- (p1 + 2 * p2) / 3
      p4 <- p3 + (p5 - p3) %*% matrix(c(0, -1, 1, 0), ncol = 2) / sqrt(3)
      new_points[(i - 1) * 4 + 2, ] <- p1
      new_points[(i - 1) * 4 + 3, ] <- p3
      new_points[(i - 1) * 4 + 4, ] <- p4
      new_points[(i - 1) * 4 + 5, ] <- p5
    }
    return(koch_curve(new_points, depth - 1))
  }
}

# Initial triangle
triangle <- matrix(c(0, sqrt(3)/2, 1, sqrt(3)/2, 0.5, 0), ncol = 2, byrow = TRUE)

# Generate Koch snowflake
koch_snowflake <- koch_curve(triangle, depth = 4)

# Plot Koch snowflake
plot(koch_snowflake, type = "l", asp = 1, xlab = "", ylab = "", main = "Koch Snowflake", col = "blue")




```


