---
title: "TIDYVERSE Assignment"
author: "angel"
date: "2024-04-07"
output: html_document
---

## Task: to create a vignette that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.

# The data set that I will be using is the current homes for sale in NY
#https://www.kaggle.com/datasets/nelgiriyewithana/new-york-housing-market

#I will be demonstrating how to use  package with readr, dplyr,ggplot2 
```{r}
library(tidyr)
library(tidyverse)
library(ggplot2)
library(dplyr)

```


#read the csv for data on current house for sales in NY
```{r}
df <- read.csv("https://raw.githubusercontent.com/Angelogallardo05/Data606-proposal/main/NY-House-Dataset.csv")
```

```{r}
head(df,3)
```

# dplyr

#Remove NA 
```{r}
df <- df %>%
  na.omit(df)%>%
  mutate(PRICE = as.numeric(PRICE))
```

#mutate to combine Queens County with Queens
```{r}
df <- df %>%
  mutate(SUBLOCALITY = case_when(
    SUBLOCALITY == "Queens" ~ "Queens County",
    TRUE ~ SUBLOCALITY  
  ))
```


#filter the data to remove outliers, and the type of home for sale to be a house in Queens
```{r}
filtered_df <- df %>%
  filter(PROPERTYSQFT != 2184.207862)%>%
  filter(TYPE == "House for sale", SUBLOCALITY == "Queens County")

# Check the filtered dataframe

```


#Summerize the average price, beds, baths and sqft. Average house for sale price in Queens is $1.1MM with about 4 beds and 2.5 baths and 1,957 sqft
```{r}
summary_stats <- filtered_df %>%
  summarise(
    count = n(),                             
    avg_price = mean(PRICE, na.rm = TRUE),   
    median_price = median(PRICE, na.rm = TRUE),  
    min_price = min(PRICE, na.rm = TRUE),     
    max_price = max(PRICE, na.rm = TRUE),     
    avg_beds = mean(BEDS, na.rm = TRUE),      
    avg_bath = mean(BATH, na.rm = TRUE),      
    avg_sqft = mean(PROPERTYSQFT, na.rm = TRUE)  
  )

# View the summary statistics
print(summary_stats)
```


# ggplot2


#use ggplot to visualize the correlation between Price and sqft, price and beds, Price and baths in queens. Seems like sqfootage and Price have a linear correlation
```{r}
ggplot(filtered_df, aes(x = PROPERTYSQFT, y = PRICE)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE) +  
  labs(x = "Property Square Footage", y = "Price", title = "Price vs. Property Square Footage") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K")) +
  theme_minimal()

ggplot(filtered_df, aes(x = BEDS, y = PRICE)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE) +  
  labs(x = "Beds", y = "Price", title = "Price vs. Beds") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K")) +
  theme_minimal()

ggplot(filtered_df, aes(x = BATH, y = PRICE)) +
  geom_point() +  
  geom_smooth(method = "lm", se = FALSE) +  
  labs(x = "Baths", y = "Price", title = "Price vs. Baths") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3, suffix = "K")) +
  theme_minimal()
```

