---
title: "TidyVerse CREATE assignment"
author: "ZIXIAN LIANG"
date: "2024-03-13"
output:
  html_document:
     toc: true
     toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Introduction


The goal of this task is to select a dataset from a web source and utilize TidyVerse packages to create a programming vignette. This vignette aims to demonstrate the use of one or more features provided by the TidyVerse package in combination with the chosen dataset.






```{r include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

## Data Acquisition


Input file:"best_selling_switch_games.csv” from

https://github.com/ZIXIANNOW/DATA607Tidyverse/blob/main/best_selling_switch_games.csv


Source:https://www.kaggle.com/datasets/kabhishm/best-selling-nintendo-switch-video-games


```{r}
url <- 'https://raw.githubusercontent.com/ZIXIANNOW/DATA607Tidyverse/main/best_selling_switch_games.csv'
database <- read_csv(url)
```



```{r}
kbl(database) %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "400px")
```


***


Use "anyNA(database)" to check for the presence of any missing values in the database.


```{r}
anyNA(database)
```

***


## Dplyr package - mutate()


Here I utilize the mutate function to create a new variable called "pub_category"


```{r}
database <- database %>%
  mutate(pub_category = case_when(
    grepl("Pokémon", publisher) ~ "The Pokémon Nintendo",
    grepl("Nintendo", publisher) ~ "Nintendo",
    TRUE ~ "Other"))
```



```{r}
kbl(database) %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "400px")
```



## Dplyr package - summarise()



Here I utilize the summarise function to aggregate data by "pub_category", calculating the total sales and counting the number of records in each category.


```{r}
datasum<- database %>%
group_by(pub_category) %>%
summarise(total_sales = sum(copies_sold),count = n())
datasum
```




## ggplot2 package - ggplot()


Here, I employ the ggplot function to generate a bar plot, representing the total sales by publisher categories.



```{r}
ggplot(datasum, aes(x = pub_category, y = total_sales, fill = pub_category)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Copy Sales by Publisher (in Millions)", x = "Publisher", y = "Total Sales") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(labels = function(x) sprintf("%.2fM", x/1000000))
```

